---
title: "RKurs - Exercise01.pdf"
author: "Venkata Subramani Renduchintala" 
date: "2 Mai 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

## Problem 1 -  Vectors
What is the output of the following commands? Try to predict the solutions before you type
in the commands. We define:
x <- c(5, 2, 1, 4); xx <- c(1, 10, 15, 18); y <- rep(1, 5)
z <- c(TRUE, FALSE, TRUE, TRUE); w <- c("Marie", "Betty", "Peter")
```{r creation of vectors, echo=TRUE}
x <- c(5, 2, 1, 4); 
xx <- c(1, 10, 15, 18); 
y <- rep(1, 5);

z <- c(TRUE, FALSE, TRUE, TRUE);
w <- c("Marie", "Betty", "Peter");

```
# Vectors - a) Basic commands with vectors
```{r basic vector operations}
sum(x);
range(x);
length(y);
sum(y)

```

# Vectors - b) c(x, y, 13)
```{r print the character vectors}
print(c(x,y,z))
c(x,y,z)
```

#Vectors - c) xx - x
c(x, 12) * y
1:6 + 1
1:9 + 1:2
```{r}
xx-x
c(x, 12) * y
1:6 + 1
1:9 + 1:2 #will give an error - non-multiple length arrays. 
```

#Vectors - 
d) x <= 2
x <= 2 & z
```{r comparison of vectors}
print (x);
print (z);
x <= 2;
x <= 2 & z
```

#Vectors - e)
substring(w, 2, 4)
paste(substring(w, 1, 2), substring(w, 5, 5), sep = "...")
```{r vector substring operations}
print(w);
substring(w, 2, 4);
paste(substring(w, 1, 2), substring(w, 5, 5), sep = "...")
```

#Vectors - f)
cbind(x, xx) #column bind 
cbind(2, 6:1, rep(c(3, 1, 4), 2), seq(1.1, 1.6, by = 0.1))
```{r Combine R objects by Rows and columns}
print(x);
print(xx);
cbind(x, xx); #column bind 
cbind(2, 6:1, rep(c(3, 1, 4), 2), seq(1.1, 1.6, by = 0.1))
```

##Problem 2 - Sequence of Numbers
Create the following sequences - Use commands rep() and seq()
#Sequence of Numbers - a) 1 2 3 4 5 6 7 8 9
```{r Numbers from 1 to 9}
seq_numbers <- seq(1:9)
print(seq_numbers)
```
#Sequence of Numbers - b) [1] "m" "w" "m" "w" "m" "w" "m" "w" "m" "w"
```{r repeat "m" "w" 5 times}
rep(c("m", "w"), 5)
```

#Sequence of Numbers - c) [1] 1 2 3 4 1 2 3 4 1 2 3 4
```{r repeat a sequence of 1 2 3 4, 3 times}
rep(seq(1:4), 3)
```

#Sequence of Numbers - d) [1] 4 4 4 3 3 3 2 2 2 1 1 1
```{r repeat a each number 3 times, in descending order}
rep.int(seq.int(4,1), c(3,3,3,3))

```

#Sequence of Numbers - e) [1] 1 2 2 3 3 3 4 4 4 4 5 5 5 5 5
```{r repeat each number in the sequence that many times}
rep(1:5, times=c(1:5))
```

#Sequence of Numbers - f) [1] 1 1 3 3 5 5 7 7 9 9 11 11
```{r repeat odd numbers twice}
rep(seq(from=1, to=11, by=2), each=2)
```


##Problem 3 - Matrices and data frames
#a) Generate the following matrices.
#    [,1] [,2] [,3] [,4]
#[1,] 1 101 201 301
#[2,] 2 102 202 302
#[3,] 3 103 203 303
#[4,] 4 104 204 304
#[5,] 5 105 205 305

# [,1] [,2] [,3]
# [1,] 5 0 0
# [2,] 0 5 0
# [3,] 0 0 5
```{r Generation of matrices}

m0 <- matrix(c(1:5, 101:105, 201:205, 301:305), nrow = 5, ncol = 4, byrow = FALSE)
m1 <- diag(x = 5, nrow = 3, ncol = 3)
print(m0)
print(m1)
```

#Matrices and data frames 
#b) Define following data frame:
# Namen Age Blind
# 1 Bob 27 TRUE
# 2 Alice 34 FALSE
# 3 Kim 21 FALSE
# 4 Julia 25 TRUE
# 5 Robert 29 TRUE
```{r data frame definition}
df0 <- data.frame(Namen = c("Bob", "Alice", "Kim", "Julia", "Robert"), 
                  Age=c(27, 34, 21, 25, 29),
                  Blind=c(TRUE, FALSE, FALSE, TRUE, TRUE))
print(df0)

```

#Matrices and data frames - c)
c) Explore the properties of your generated objects. Which class of R-objects do they belong to? How are they structured?
Hint: Use the functions class(), dim(), str(), summary().
```{r Properties of generated objects}
class(m0) #give typee of object
dim(m0) #dimension of an object
str(m0) #Structure
summary(m0) #summary per column

```

